---
title: "Startercode"
author: "Rikke Kjems & Pernille Ryder"
date: "4/28/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Setting up working directory and loading data
```{r Setting up working directory and loading data}
#Setting working directory

#Pernille
#setwd("/Users/pernille/Desktop/DS_Exam")

#Rikke
#setwd("~/Library/Mobile Documents/com~apple~CloudDocs/Cognitive Science/10. Semester/Data Science/Exam")

#Loading data
data <- read.csv("weatherAUS.csv", header = TRUE, sep = ",", stringsAsFactors = TRUE)
```



#Loading libraries
```{r Loading libraries}
#Loading libraries

library(ggplot2)
library(dplyr)
library(tsibble)
library(feasts)
library(fpp3)
library(data.table)
library(tidyverse)
library(plotly)
library(caret)
library(magrittr)
library(corrplot)
library(caTools)
library(pROC)
library(ggthemes)
library(cowplot)
library(dplyr)
library(tsibble)
library(plyr)
library(ggplot2)
library(hydroTSM)
library(lubridate)
library(fable.prophet)
library(features)
library(forecast)

```



#Data Preprocessing
```{r cars}
set.seed(1023) 
#Making a subset without Date, Location, RainToday and Raintoday. as these variables are essential.
weather_data <- subset(data, select = -c(Date, Location, RainToday, RainTomorrow))
colnames(weather_data) #printing the column names
summary(weather_data)
```

```{r Histograms of skewness} 
#Check the skewness of data
hist(weather_data$MaxTemp)
hist(weather_data$MinTemp)
hist(weather_data$Evaporation)
hist(weather_data$Sunshine)
hist(weather_data$WindGustSpeed)
hist(weather_data$WindSpeed9am)
hist(weather_data$WindSpeed3pm)
hist(weather_data$Humidity9am)
hist(weather_data$Humidity3pm)
hist(weather_data$Temp9am)
hist(weather_data$Temp3pm)
hist(weather_data$Pressure9am)
hist(weather_data$Pressure3pm)

```

```{r feature extraction -> chisquare values}
#2.Feature Extraction
#Get the categorical variables
#2.1Chi-Square to check whether the variables are dependent on RainTomorrow

#Selecting the factorial variables
factor_vars1 <- names(which(sapply(weather_data, class) == "factor"))
factor_vars1

#Raintomorrow is removed
factor_vars1 <- setdiff(factor_vars1, "Rainfall")
factor_vars1

#chi sqaured test
chisq_test_res <- lapply(factor_vars1, function(x) { 
  chisq.test(weather_data[,x], weather_data[, "Rainfall"], simulate.p.value = TRUE)
})
names(chisq_test_res) <- factor_vars1
chisq_test_res

```

```{r Removing "WindGustDir" "WindDir9am" & "WindDir3pm"}

#Remove Categorical variables from data set
numeric_vars <- subset(weather_data, select = -c(WindDir9am, WindDir3pm, WindGustDir))
colnames(numeric_vars)

```

```{r Correlation plot}
#Feature Extraction for numeric variables

#Ensuring no NA's in order to make correlation plot
numeric_vars <- numeric_vars[complete.cases(numeric_vars),] 

#Calculating correlation values for each variable 
numeric_vars_cor <- cor(numeric_vars)

#correlation plot
corrplot(numeric_vars_cor, method = "pie")

```

```{r Calculating amount of NA in each column}
#Calculating how many NA in % in each column
apply(data, 2, function(col)sum(is.na(col))/length(col))*100

#Every variable with more than 15 % of NA will be excluded
#The three categorical variables will also be excluded based on the pearson correlation calculation done above.
```

```{r Making subset with only the variables we want}
#Making subset data with only the variables we want. We exclude every variable with more than 15 % NA

final_data <- subset(data, select = -c(Evaporation, Sunshine, Cloud9am, Cloud3pm, WindGustDir, WindDir9am, WindDir3pm, Temp9am, Temp3pm, MinTemp, Pressure3pm, Humidity3pm))

final_data <- final_data[complete.cases(final_data),]
```

```{r Creting time columns with different time configurations}

# Now we create different columns for different time configurations

# Making a variable for Year
final_data  <- mutate(final_data , Year = year(Date))

final_data$Year <- factor(final_data$Year)


```

```{r plotting rainfull over time for all locations}
#Plotting rainfall over time for each location
ggplot() + geom_line(data = final_data, aes(x = Date, y = Rainfall, color = Location), group=1)

```


```{r Making day column}
#New column with 0-365 for each year

date <- ymd(final_data$Date)

days <- yday(date) - 1 # so Jan 1 = day 0 

final_data  <- mutate(final_data , Day = days) # Adding list 'days'-list to DF

final_data$Day <- factor(final_data$Day)
```

```{r Making a tsibble}
tsb <- final_data %>%
  as_tibble()

Rain <- tsb %>%
  mutate(Date = as.Date(Date)) %>%
  as_tsibble(index = Date, key = 'Location')
```


```{r Plot of rain across all years}

ggplot() + geom_line(data = Rain, aes(x = Day, y = Rainfall), group=1) +
  facet_wrap(~Year)
```

#Checking gaps in data

```{r checking for gaps in data}
#has_gaps(Rain, .full = TRUE)
#scan_gaps(Rain)

rain_gaps <- Rain %>% 
  count_gaps(.full = TRUE)
rain_gaps
```

```{r visualising gaps}
ggplot(rain_gaps, aes(x = Location, colour = Location)) +
  geom_linerange(aes(ymin = .from, ymax = .to)) +
  geom_point(aes(y = .from)) +
  geom_point(aes(y = .to)) +
  coord_flip() +
  theme(legend.position = "bottom")

#Plot shows huge gab in data from Sydney from 2009-2011. We choose to exclude all data from before 2011.
#Many individual gabs and a huge long one i CoffsHarbour. This location will be excluded. 
#We move on with data from 2011-2017 from location Wollongong, WaggaWagga, Sydney, Albury
```

```{r Removing data from before 2011 and exclude CoffsHarbour}
#Removing data from before 2011 and exclude CoffsHarbour:
#Rain <- Rain%>%filter(Year%in%c(2011:2017)) 

#Rain <- Rain%>%filter(Location%in%c("Wollongong", "WaggaWagga", "Sydney", "Albury")) 

new_rain_gaps <- Rain %>% 
  count_gaps(.full = TRUE)
new_rain_gaps
ggplot(new_rain_gaps, aes(x = Location, colour = Location)) +
  geom_linerange(aes(ymin = .from, ymax = .to)) +
  geom_point(aes(y = .from)) +
  geom_point(aes(y = .to)) +
  coord_flip() +
  theme(legend.position = "bottom")

```

```{r Handle implicit missingness}

#Filling out the gaps in the Date column, returns NA in all other columns where there is gap
rain_full <- Rain %>% 
  fill_gaps(.full = TRUE)
rain_full

```

```{r Filling out gaps in Location, RainToday and RainTomorrow}
#Filling out gaps in Location, RainToday and RainTomorrow with previous values

rain_full <- rain_full %>% 
  group_by_key() %>% 
  fill_gaps() %>%
  tidyr::fill(Location, .direction = "down")

rain_full <- rain_full %>% 
  group_by_key() %>% 
  fill_gaps() %>%
  tidyr::fill(RainToday, .direction = "down")

rain_full <- rain_full %>% 
  group_by_key() %>% 
  fill_gaps() %>%
  tidyr::fill(RainTomorrow, .direction = "down")

```

```{r Filling out gaps in remaining variables}

#Filling out gaps by using group_by_key and replace.na for finding the mean for each key=Location

rain_done <- rain_full %>% 
  as_tsibble(index=Date, key=Location) %>% 
  group_by_key() %>% 
  mutate(MaxTemp = tidyr::replace_na(MaxTemp, mean(MaxTemp, na.rm=T)))%>%
  mutate(WindGustSpeed = tidyr::replace_na(WindGustSpeed, mean(WindGustSpeed, na.rm=T)))%>%
  mutate(WindSpeed9am = tidyr::replace_na(WindSpeed9am, mean(WindSpeed9am, na.rm=T)))%>%
  mutate(WindSpeed3pm = tidyr::replace_na(WindSpeed3pm, mean(WindSpeed3pm, na.rm=T)))%>%
  mutate(Humidity9am = tidyr::replace_na(Humidity9am, mean(Humidity9am, na.rm=T)))%>%
  mutate(Pressure9am = tidyr::replace_na(Pressure9am, mean(Pressure9am, na.rm=T)))%>%
  mutate(Rainfall = tidyr::replace_na(Rainfall, mean(Rainfall, na.rm=T)))

```

```{r Filling out gaps in time columns}
#Filling out the gaps in Day column by simply making the columns again

#New column with 0-365 for each year

date <- ymd(rain_done$Date)

days <- yday(date) - 1 # so Jan 1 = day 0 

rain_done  <- mutate(rain_done , Day = days) # Adding list 'days'-list to out DF

rain_done$Day <- factor(rain_done$Day)

# Making a column for Year
rain_done  <- mutate(rain_done , Year = year(Date))

rain_done$Year <- factor(rain_done$Year)
```

```{r Making rain_done into a tsibble}
#Making rain_done into a tsibble
rain_done <- rain_done %>%
  mutate(Date = as.Date(Date)) %>%
  as_tsibble(index = Date, key = 'Location')
rain_done
```

```{r Making sure that all gaps are filled}
#Making sure that all gaps are filled
has_gaps(rain_done, .full = TRUE)
scan_gaps(rain_done)
rain_gaps <- rain_done %>% 
  count_gaps(.full = TRUE)
rain_gaps

ggplot(rain_gaps, aes(x = Location, colour = Location)) +
  geom_linerange(aes(ymin = .from, ymax = .to)) +
  geom_point(aes(y = .from)) +
  geom_point(aes(y = .to)) +
  coord_flip() +
  theme(legend.position = "bottom")

#Plot is empty because of no gaps
```


#Visualising some data
```{r Rainfall over time}
#Plot Time Series Data
rain_done%>%
  #filter(Location=="Sydney")%>%
autoplot(Rainfall) + ylab("Rainfall") + xlab("Day")+
  #scale_x_date(date_labels = '%Y', breaks = '1 year') +
  theme_bw() + ggtitle("Rainfall in Austrlia")
  #facet_wrap(~Year)
  

#ggplot() + geom_line(data = rain_done, aes(x = Day, y = Rainfall), group=1) +
 # facet_wrap(~Year)

```

#ALbury subset
```{r}

albury = rain_done %>%
  filter(Location =='Albury')

albury <- subset(albury, select = -c(Location, RainToday, RainTomorrow))

```

```{r Albury}

albury %>%
  autoplot(Rainfall)

```

```{r Albury}
ggplot() + geom_line(data = albury, aes(x = Day, y = Rainfall), group=1) +
  facet_wrap(~Year)

#The facet wrap plot indicates that 2007, 2008, and 2017 is not full years, and they are therefore excluded in our albury data set. 

`````


````{r data filtering}
#on basis of the facet wrap plot
albury<- albury[albury$Year %in% c('2009','2010','2011','2012','2013','2014','2015','2016'), ]

```````


```{r new facet wrap}

ggplot() + geom_line(data = albury, aes(x = Day, y = Rainfall), group=1) +
  facet_wrap(~Year)

`````

#Checking for seasonality & trends
```{r gg_season plot}
albury %>%
  gg_season(Rainfall, labels = "both") +
  labs(y = "Rainfall",
       title = "Seasonal plot: Rainfall over time in Albury")

#filter out one city
```

```{r ACF Rainfall}
albury %>%
  ACF(Rainfall, lag_max = 365)

albury%>%
  ACF(Rainfall, lag_max=365)%>%
  autoplot() + labs(title="Rainfall in Albury")

```

# Albury
```{r STL plot for Rainfall}
dcmp <- albury %>%
  model(stl = STL(Rainfall~season(window=8))) #Because we have eight seasons/years
components(dcmp)%>%
  autoplot()

```


#Arima model

````{r}
albury %>%
  model(ARIMA(Rainfall)) %>% 
          forecast(h = "2 years") %>% 
          autoplot(albury) + labs(title="ARIMA model forecasting 2 years")
``````

#Check Arima model
```{r}

albury%>%
  model(ARIMA(Rainfall)) %>%
   gg_tsresiduals() + labs(title="ARIMA model performance")

```


# Mean model 
````{r}
albury %>%
  model(MEAN(Rainfall)) %>% 
          forecast(h = "2 years") %>% 
          autoplot(albury) + labs(title="Mean model forecasting 2 years")
``````

#Check mean  model
```{r}

albury%>%
  model(MEAN(Rainfall)) %>% 
   gg_tsresiduals() + labs(title="Mean model performance")

```


#TSLM
````{r}
albury %>%
  model(TSLM(Rainfall)) %>% 
          forecast(h = "2 years") %>% 
          autoplot(albury) + labs(title='TSLM model forecasting 2 years')
``````

#Check TSLM  model
```{r}

albury%>%
  model(TSLM(Rainfall)) %>% 
   gg_tsresiduals() + labs(title='TSLM model performance')

```

#Vector Autoregression
```{r Vector autoregression model}

fit <- albury %>%
  model(VAR(vars(Rainfall)))

glance(fit)


fit %>%
  augment() %>%
  ACF(.innov) %>%
  autoplot() + labs(title='VAR model forecasting 2 years')


#Lag 29 is significant as the only one. Meaning this model is doing okay.
`````

```{r Vector autoregression model plot}

fit %>%
  forecast(h = '2 years') %>%
  autoplot(albury) + labs(title='VAR model forecasting 2 years')

````
```{r}

albury%>%
  model(VAR(vars(Rainfall))) %>% 
   gg_tsresiduals() + labs(title='VAR model performance')

````


#Test and traning set -fpp3 5.8

```{r}

#albury_test<- albury[albury$Year %in% c('2015','2016'), ]

albury_train<- albury[albury$Year %in% c('2009','2010','2011','2012','2013','2014'), ]


`````


```{r}

albury_fit <- albury_train %>%
  model(
    mean = MEAN(Rainfall),
    tslm = TSLM(Rainfall ~ trend()),
    arima = ARIMA(Rainfall),
  )

``````

```{r}

albury_fc <- albury_fit %>%
  forecast(h = '2 years')
  

````


````{r}

#Checking accuracy
albury_fc%>%
  accuracy(albury)%>%
  arrange(MASE)


````

```{r}
#Plot test/train
albury_train %>%
    model(
    mean = MEAN(Rainfall),
    tslm = TSLM(Rainfall ~ trend()),
    arima = ARIMA(Rainfall),
  ) %>%
  forecast(h = '2 years') %>%
  autoplot(albury) + labs(title='Mean, TSLM, and ARIMA fitted on train data')

`````


#VAR model fitted to train data 

```{r}
var_fit <- albury_train %>%
  model(VAR(vars(Rainfall)))

````

```{r}

var_fc <- var_fit %>%
  forecast(h = '2 years')

````
  
```{r}
var_fc %>%
accuracy(albury)%>%
arrange(MASE)

````

```{r}

albury_train %>%
  model(VAR(vars(Rainfall))) %>%
 forecast(h = '2 years') %>%
  autoplot(albury) + labs(title = 'VAR model fitted on train data')
````

#Differencing albury data (removing trend and seasonality)
```{r}

#After one differencing the seasonality is removed
albury %>%
  gg_tsdisplay(difference(Rainfall, 12), #Annual change in rainfall, 365 if daily.
               plot_type='partial', lag=8)

```
#Vector Autoregression after difference 
```{r Vector autoregression model}

fit <- albury %>%
  model(VAR(vars(Rainfall)))

glance(fit)


fit %>%
  augment() %>%
  ACF(.innov) %>%
  autoplot() + labs(title='VAR model performance after difference')

`````

````{r}

fit %>%
  forecast(h = '2 years') %>%
  autoplot(albury) + labs(title='VAR model forecasting 2 years')
`````


```{r}


albury%>%
  model(VAR(vars(Rainfall))) %>% 
   gg_tsresiduals() + labs(title='VAR model performance')

`````

# Mean, tslm and arima after difference
```{r}
albury %>%
    model(
    mean = MEAN(Rainfall),
    tslm = TSLM(Rainfall ~ trend()),
    arima = ARIMA(Rainfall)
  ) %>%
  forecast(h='2 years') %>%
  autoplot(albury) + labs(title='Mean, TSLM, ARIMA forecasting after differencing')
`````

#check models after difference
```{r}

albury%>%
  model(TSLM(Rainfall)) %>% 
   gg_tsresiduals()

`````

```{r}

albury%>%
  model(MEAN(Rainfall)) %>% 
   gg_tsresiduals()

`````

```{r}

albury%>%
  model(ARIMA(Rainfall)) %>% 
   gg_tsresiduals()

`````




